---
title: "Music Motor Manuscript"
output:
  pdf_document:
      latex_engine: xelatex
  html_document:
    df_print: paged
  word_document: default
date: "2026-04-09"
bibliography: references.bib
editor_options:
  markdown:
    wrap: sentence
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

Prior research suggests that music-evoked emotions may drive movement, thereby reinforcing emotional experience through engagement of the motor system.
This relationship has been linked to activity in the primary motor cortex (M1); however, its causal role remains unclear.
The present within-subjects study examined whether inhibiting M1 using continuous theta-burst stimulation alters perceived valence and arousal ratings of music.
Contrary to our hypothesis, no significant effects of M1 inhibition on valence or arousal ratings were observed relative to sham stimulation.
Exploratory analyses examining musical sophistication, as measured by the Goldsmith Musical Sophistication Index (Gold-MSI), revealed that arousal ratings were significantly associated with the Active Engagement and Emotional Engagement subscales.
In contrast, no significant relationships were observed between Gold-MSI scores and valence ratings.
Gold-MSI scores did not moderate the effects of stimulation condition.
These findings suggest that M1 may not play a direct causal role in emotional responses to music, highlighting the need to investigate other motor-related regions.

Keywords: primary motor cortex (M1); continuous theta-burst stimulation (cTBS); valence; arousal; music.

```         
```

## Introduction

Emotion and movement are tightly interconnected, as physical action is not only an expression of emotion but also a contributor to its production [@shafir2013].
This idea is rooted in the James–Lange theory, which proposes that emotions arise as a response to bodily changes (e.g., fear emerges from physiological reactions such as trembling and increased heart rate).
Damasio’s somatic marker hypothesis further extends this framework by suggesting that the brain stores bodily states as memories and links them to emotional experiences.
As a result, when individuals perform, observe, or imagine a movement, the brain reactivates these bodily states to generate emotion in the current context.
Consistent with these theories, Shafir et al. (2013) demonstrated that executing, observing, and imagining body movements associated with specific emotions (e.g., happiness, sadness, fear) increased the corresponding affective state.
The cross-modal bias theory further extends these frameworks by proposing that emotional information from one sensory modality, such as music, influences how emotions are experienced in another modality, such as body movement [@christensen2014].
Christensen et al. (2014) demonstrated that emotional experience can be characterized along two dimensions: valence, which refers to how positive or negative an emotion is, and arousal, which refers to the level of energy.
Their results showed that subjective emotional ratings were primarily driven by valence; ergo, music influenced whether dance movements were perceived as more positive or negative.
In contrast, physiological responses were modulated by arousal congruency, with the strongest bodily responses occurring when low-arousal movements were paired with sad music and high-arousal movements were paired with happy music.
Yet, despite growing evidence of an association between movement and emotion, only recently has research begun to employ noninvasive neurophysiological techniques to investigate this correlation.
Doruk et al. (2016) helped address this gap in the literature by providing evidence that specific neurophysiological markers of the motor system are associated with post-stroke emotional disturbances, such as anxiety and depression.
Using electroencephalography (EEG) and transcranial magnetic stimulation (TMS), the study demonstrated that emotional disturbance was not solely attributable to limbic dysfunction, but also reflected altered motor system dynamics.
EEG revealed reduced interhemispheric coherence and frontal–central connectivity, while TMS indicated increased intracortical facilitation and higher motor thresholds in the unlesioned hemisphere.
These neurophysiological markers were not associated with motor impairment severity; however, they were associated with mood and emotional control.
Collectively, these findings indicate that motor and emotional disturbances following stroke are supported by distinct yet interacting neural mechanisms, rather than emotional disturbance being a direct consequence of motor impairment.
Selosse et al. (2025) sought to further evaluate the relationship between movement and emotion by isolating components of the motor system, specifically examining the relationship between the primary motor cortex (M1) and emotion.
The researchers used functional magnetic resonance imaging (fMRI) to examine emotional prosody production while participants produced pseudowords in angry, happy, and neutral tones under three conditions: normal speech, whispering, and silent articulation.
The authors found greater M1 activation during normally voiced emotional speech compared to whispering or silent articulation.
This pattern suggests that emotional prosody relies on a broader neural network that includes motor regions, with M1 supporting the execution of the vocal movements required to express emotion.
Thus, emotional voice production can be understood as an embodied process in which M1 contributes to the physical expression of emotion through vocal tract movement and speech.
Xia et al. (2021) further examined the relationship between the M1 and emotion by distinguishing the associations between emotional processing and activity in the left and right M1.
The study employed TMS to record motor-evoked potentials (MEPs) from both the left and right M1 while right-handed participants were instructed to move a manikin toward positively valenced images and away from negatively valenced images using keyboard presses.
The results showed that the left M1 exhibited larger MEP amplitudes, indicating increased excitability, when participants performed compatible behaviors (i.e., approaching positive stimuli and avoiding negative stimuli).
In contrast, the right M1 showed reduced excitability during these compatible behaviors.
These effects emerged prior to physical movement; therefore, during stimulus processing.
Together, these findings suggest that emotional valence modulates motor preparation, supporting a functional association between emotion processing and contralateral M1 activity.
Despite growing evidence linking M1 activity to emotional processing, most research has established only correlational relationships.
Establishing causality requires noninvasive neuromodulation techniques, such as continuous theta-burst stimulation (cTBS), which can inhibit cortical regions to assess their functional contributions.
Prior work has demonstrated that cTBS applied to M1 can be used to test for causal involvement by examining behavioral consequences following inhibition (e.g., Kraeutner et al., 2017).
However, little research has directly investigated whether inhibiting M1 alters self-reported emotions, leaving the causal role of M1 in emotion largely unresolved.
Hence, the present study aims to address this gap by using cTBS to inhibit the contralateral M1 and examining its effects on emotion production in response to pre-rated music clips varying in valence and arousal.
In a within-subjects design, right-handed participants completed two sessions: active cTBS (experimental) and sham (control).
After stimulation, participants listened to instrumental music clips and rated each clip’s valence and arousal.
Participant musical sophistication was also measured using a questionnaire.
We hypothesize that M1 inhibition will yield more neutral valence and arousal ratings than sham, reflecting reduced M1 involvement in emotion processing.
Additionally, if participants possess greater musical sophistication, they may show more polarized ratings under M1 inhibition than those with lower musical sophistication, suggesting alternative emotion-processing mechanisms less dependent on M1.

```         
```

## Methods

**Participants**

Study participants (N = 29) were recruited through the SONA online platform and poster advertisement.
Eligibility was restricted to right-handed individuals who were able to understand English and had normal or corrected-to-normal vision.
Participants were excluded if they self-reported any neurological conditions or impairments that would affect motor function, or if they had any contraindications to TMS as identified by a standard screening form (Rossi et al., 2009).
For the final analysis, N = 22 university students were included, and 14 identified as female.
Participants had a mean age of 22.73 years (SD = 4.7, range = 19–38) and a mean General Musical Sophistication score of 74.68 (SD = 16.45, range = 35–114).
Each participant received either \$10 CAD per session or SONA course credit (2 credits for Session 1 and 1.5 credits for Session 2).
Data collection took place between July 15, 2025, and February 26, 2026.

**Materials**

**Transcranial Magnetic Stimulation (TMS) Screening Form**

To ensure participant eligibility and safety during transcranial magnetic stimulation (TMS), the TMS Screening Form was administered (Rossi et al., 2009).
The questionnaire includes 10 binary items (yes/no) assessing medical history; a “yes” response to any item results in exclusion.
Participants were also required to report current medications, the presence of metal in the body, and any prior experience with TMS or magnetic resonance imaging (MRI).
Any medications, implanted metal, or history indicating a contraindication or adverse response to TMS or MRI led to exclusion.

**Kinaesthetic and Visual Imagery Questionnaire (KVIQ)**

The Kinaesthetic and Visual Imagery Questionnaire (KVIQ) was used to assess participants’ ability to perform kinaesthetic and visual motor imagery (Malouin et al., 2007).
Participants physically performed five body movements and then imagined each movement using visual imagery, followed by kinaesthetic imagery.
Participants rated the clarity and intensity of the imagined movement on a 5-point scale, with 1 indicating the weakest imagery and 5 indicating the strongest.

**Goldsmith Musical Sophistication Index (Gold-MSI)**

To assess musical sophistication, participants completed the Goldsmiths Musical Sophistication Index (Gold-MSI; Müllensiefen et al., 2014).
This questionnaire evaluates six dimensions: active engagement, perceptual ability, musical training, emotional engagement, singing ability, and general musical sophistication.
Participants responded to items using a 7-point Likert scale, with 1 indicating strong disagreement and 7 indicating strong agreement.
The questionnaire was also used to collect descriptive information, including participants’ age, sex, and education level.

**Transcranial Magnetic Stimulation (TMS)**

Participants underwent transcranial magnetic stimulation (TMS)– a non-invasive technique that uses brief magnetic pulses to generate electrical currents in the brain and will be used in accordance with well-established procedures (Rossi et al. 2009; Huang et al. 2005; Wischnewski et al. 2015).
Stimulation output intensity has been individualized using established protocols (Kleim et al., 2007; Rossi et al., 2009).
Neuronavigation (Brainsight 2™; Rogue Research Inc.) provided focal, real-time localization of the TMS coil with millimetric accuracy by optically tracking coil position relative to digitized anatomical landmarks co-registered to a template brain.
This system enabled precise targeting of cortical regions.
To determine stimulation intensity and the motor “hotspot,” resting motor threshold (RMT) was assessed as an index of cortical excitability.
RMT is defined as the minimum stimulator output that elicits visible muscle twitches and motor-evoked potentials ≥50 μV.
This was recorded using surface electromyography (EMG) over the left first dorsal interosseous (FDI) in at least 5 of 10 trials.
Baseline motor excitability was assessed by recording 16 motor evoked potentials (MEPs) at rest, 16 during kinaesthetic motor imagery (KMI), and 16 during visual motor imagery (VMI) of a simple finger abduction.
TMS was delivered at 120% of resting motor threshold (RMT) (Bruton et al., 2020; Stinear et al., 2007; Wright et al., 2014).
Participants received continuous theta-burst stimulation (cTBS) over the contralateral M1 in accordance with established guidelines (Huang et al., 2005; Rossi et al., 2009, 2020; Wischnewski et al., 2015).
The cTBS protocol consisted of bursts of three pulses at 50 Hz, repeated every 200 ms for 40 s (600 pulses) at 80% RMT, delivered using a Magstim SuperRapid2 Plus-1 with a cooled figure-of-eight coil.
Sham cTBS was delivered to the vertex at 20% RMT for an equivalent duration using a validated protocol that mimics the sensory experience of TMS without producing physiological effects (Kraeutner et al., 2016, 2017).
All stimulation parameters were within established safety standards.
Following cTBS or sham stimulation, MEPs were recorded under the same conditions to confirm M1 inhibition.

**Active Listening Task**

Participants’ emotion production was assessed by completing an active listening task, which followed previously reported procedures (Eerola & Vuoskoski, 2011).
Participants were randomly assigned to listen to one of two sets of 32 pre-rated instrumental music clips, with clip order randomized across sessions and participants.
The clips were balanced for valence and arousal and were selected from an established set of 110 clips (openly available at <https://osf.io/p6vkg/wiki/home/>).
During the task, participants were seated in a chair with their dominant hand resting on their lap and their non-dominant hand placed on a keypad.
They listened to the clips through headphones and, after each clip, used the keypad to rate their emotional responses on 9-point valence and arousal scales.
Procedure Participants were informed that participation was voluntary and that they could withdraw from the study at any time without penalty or loss of compensation.
All participants completed a TMS safety screening form to confirm eligibility and provided written informed consent before participation.
See Figure 1 for a schematic overview of the study procedure.
The study consisted of two sessions separated by a minimum of 24 hours.
In Session 1, participants completed the KVIQ, the Motor Experience Questionnaire, and the Gold-MSI.
Participants were then seated comfortably, and surface EMG electrodes were placed on the left FDI to record muscle activity.
Using neuronavigation, the contralateral M1 was localized, and RMT was determined.
Participants subsequently received 16 single-pulse TMS stimulations at rest, 16 during KMI, and 16 during VMI.
The motor imagery task consisted of imagining abduction of the left index finger.
Imagery trials were cued using the software PsychoPy; participants were instructed to close their eyes and begin imagining upon hearing the auditory cue “Go” and to stop imagining upon hearing “Rest.” Each imagery period lasted 10 seconds, during which two TMS pulses were delivered.
Participants were permitted to rest between trials and initiated each subsequent trial by pressing a keyboard key.
Following baseline stimulation, participants received either active or sham cTBS.
Earplugs were provided to minimize auditory discomfort.
All participants received both active and sham cTBS across two sessions, with the order randomized.
Participants were informed that they would receive both conditions, but were blinded to the condition administered in each session.
After cTBS, participants were instructed to refrain from moving their left hand for 40 minutes.
A 10-minute rest period followed cTBS, after which participants again received 16 TMS pulses at rest, 16 during KMI, and 16 during VMI.
The order of KMI and VMI conditions was counterbalanced across sessions.
Participants then completed an active listening task.
Using a PsychoPy program, participants listened to 32 instrumental music clips through headphones.
After each clip, they rated perceived valence and arousal on a 9-point scale using keyboard responses.
The order of valence and arousal ratings was counterbalanced across sessions.
All post-cTBS tasks were completed within 40 minutes, and no data were collected beyond this period.
At the conclusion of each session, participants received compensation in the form of cash or SONA research credit.
See Figure 1 for a visual of the study methodology.

**Statistical analyses**

All statistical analyses were conducted using RStudio.
Participants were randomly assigned to receive cTBS in session 1 or session 2.
The order of motor imagery task performance—KMI and VMI—as well as the rating of valence and arousal, was counterbalanced across sessions.
Each participant contributed eight observations per session, corresponding to ratings of the four pre-defined categories of valence and arousal: high valence–low arousal, low valence–high arousal, high valence–high arousal, and low valence–low arousal.
Participants provided ratings in both the sham control and active cTBS conditions.
Univariate analysis of variance (ANOVA) were conducted to assess the effects of active versus sham cTBS on valence and arousal ratings across the four clip categories.
Linear mixed-effects models (LMEs) were conducted to examine whether Gold-MSI ratings mediated the relationship between M1 inhibition and perceived valence and arousal ratings.
Music clips were selected from a validated set of 110 clips.
Four categories were created: 16 clips with the largest difference between high valence and low arousal, 16 clips with the largest difference between low valence and high arousal, 16 clips with the highest valence and arousal, and 16 clips with the lowest valence and arousal.
These four groups were then randomly divided into two sets of eight clips per category, yielding two sets of 32 music clips.
Participants were randomly assigned to listen to one of the two sets.
The presentation order of the clips was randomized within each participant’s session and across participants.

```         
```

## Results

Following the criteria above, seven participants were excluded due to inability to establish a reliable motor threshold (n = 4), issues during the session (n = 1), use of medication contraindicated for TMS (n = 1), and participant failure to attend the session (n = 1).
Hence, data from 22 participants were included in the final analysis.

**Active Listening Task**

To test the hypothesis that inhibition of M1 would neutralize perceived valence and arousal ratings, two separate univariate analyses of variance (ANOVAs) were conducted.
For both mean valence and mean arousal ratings, there was no significant main effect of stimulation condition (see Figures 2 & 3).
However, a significant main effect of stimulus type was observed for both valence (p \< .001) and arousal (p \< .001) using Greenhouse–Geisser corrections (see Figures 2 & 3).
This indicates that ratings differed across the four stimulus categories.
Effect size analyses further indicated that the stimulation condition had minimal influence on both valence and arousal ratings.
For valence, Cohen’s d values were small or negligible across stimulus categories: high valence–low arousal (d = 0.37), indicating a small-to-moderate effect with higher ratings in the active condition; low valence–low arousal (d = 0.23), indicating a small effect with higher ratings in the active condition; low valence–high arousal (d = −0.15), indicating a very small effect with lower ratings in the active condition; and high valence–high arousal (d = 0.01), indicating no meaningful effect.
Similarly, for arousal ratings, effect sizes were small or negligible: high valence–low arousal (d = −0.23), indicating a small effect with lower ratings in the active condition; low valence–high arousal (d = 0.15), indicating a very small effect with higher ratings in the active condition; low valence–low arousal (d = 0.07), indicating no meaningful effect; and high valence–high arousal (d = 0.02), also indicating no meaningful effect.

**General Musical Sophistication**

To examine whether musical sophistication was associated with perceived valence and arousal ratings across stimulus types, linear mixed-effects models (LMEs) were conducted, with participant included as a random intercept.
Gold-MSI scores (Active Engagement, Perceptual Ability, Musical Training, Emotional Engagement, Singing Ability, and General Musical Sophistication) were entered as fixed effects, along with stimulus type.
The model predicting mean valence ratings revealed no significant main effects of Gold-MSI scores on any subscale, indicating that individual differences in musical sophistication did not significantly predict mean valence ratings across stimulus categories.
The model predicting mean arousal ratings revealed significant effects of the Gold-MSI subscales Active Engagement (p = .004) and Emotional Engagement (p = .005; see Figures 4 & 5).
In contrast, the remaining subscales (Perceptual Ability, Musical Training, Singing Ability, and General Musical Sophistication) were not significantly associated with mean arousal ratings.
Additional LMEs examining both mean valence and arousal ratings across stimulus type and stimulation condition revealed no significant main effects involving Gold-MSI scores across the six subscales.

```         
```

## Discussion

We hypothesized that inhibition of M1 would lead to more neutral valence and arousal ratings compared to sham stimulation, reflecting a reduced role of M1 in emotion processing.
Additionally, we predicted that individuals with greater musical sophistication would exhibit more polarized emotional ratings under M1 inhibition, suggesting reliance on alternative emotion-processing mechanisms.
Contrary to these predictions, stimulation condition did not significantly influence either valence or arousal ratings, and effect size analyses indicated only minimal differences between active and sham conditions.
Instead, emotional ratings were primarily driven by stimulus type, with clear differences observed across the four stimulus categories.
Furthermore, musical sophistication did not moderate the effects of M1 inhibition on emotional ratings.
Although exploratory analyses revealed that Active Engagement and Emotional Engagement were associated with higher arousal ratings, no such relationships were observed for valence.
Together, these findings do not support a causal role of M1 in shaping emotional responses to music, nor do they indicate that musical sophistication alters this relationship.
One possible explanation for the absence of M1 effects is that the active listening task may not have strongly engaged motor-related processes (Lee & Siegle, 2012).
Prior research suggests that different types of emotional evaluations recruit distinct neural networks.
For example, stimulus-focused evaluations, such as rating the emotional content of a stimulus, primarily engage sensory processing regions and prefrontal areas involved in evaluation, including the ventrolateral prefrontal cortex.
In contrast, evaluating one’s own emotional state recruits regions such as the insula and anterior cingulate cortex, while evaluating others’ emotions engages social-cognitive regions such as the superior temporal sulcus and temporoparietal junction.
Since participants in the present study rated the perceived valence and arousal of musical stimuli, their responses may have relied more on perceptual and evaluative networks than on motor system involvement.
This may help explain why inhibiting M1 did not significantly alter emotional ratings.
The lack of M1 involvement in the present study may also be explained by the task's limited social and interactive context (Antonelli et al., 2025).
Inter-neural synchrony (INS), which reflects coordinated brain activity across individuals, may emerge most strongly during real-time, reciprocal social interactions where individuals continuously exchange emotional and behavioral cues.
In contrast, when participants engage with non-social or computer-mediated stimuli, such as rating music on a device, the interaction lacks the dynamic feedback loop necessary to support strong interpersonal coupling.
This, in turn, may reduce INS.
Since motor regions such as M1 are often implicated in action coordination and interpersonal synchrony, reduced INS in a non-interactive setting may limit recruitment of these regions.
Consequently, the absence of a meaningful social context in this study may have attenuated M1 involvement; thereby, inhibitory stimulation did not significantly alter participants’ emotional ratings of the music.
Another possible explanation for why inhibiting the M1 did not alter participants’ valence or arousal ratings of music clips is that M1 may not be central in generating affective evaluations of music (Putkinen et al., 2021).
Although the motor cortex contains emotion-specific activity patterns and supports embodied aspects of music perception (e.g., movement), fundamental affective dimensions such as valence and arousal are primarily mediated by auditory cortical processing and by broader limbic and paralimbic systems.
As a result, any potential influence of musical sophistication on M1-related emotional processing may not be detectable because there are no observable differences in valence or arousal experiences relative to sham.
Although exploratory analyses showed that Active Engagement and Emotional Engagement were associated with higher arousal ratings, no such relationships were observed for valence.
One possible explanation is that participants who are more emotionally and actively engaged with music may experience stronger empathic or embodied responses during the active listening task (Chapin et al., 2010).
Music may evoke emotion by prompting participants to internally simulate the tempo and its expression, making the experience feel more intense.
Individuals who frequently engage with music or are more emotionally attuned may be more sensitive to these expressive cues, leading to heightened arousal.
In contrast, valence appears to be more subjective and less consistently tied to these embodied processes, which may explain why Active and Emotional Engagement did not influence whether emotions were perceived as positive or negative.
Future research should move beyond numerical valence and arousal scales, as they are limited in their ability to represent emotion and can be interpreted differently by participants, introducing variability in the data.
Future studies could include discrete emotion categories such as happiness, sadness, fear, anger, and tenderness, alongside dimensional measures, to better capture the specific emotional aspects of music (Eerola & Vuoskoski, 2011).
Additionally, distinguishing among types of arousal, such as energy and tension arousal, may provide more detail about how emotions are experienced.
Including measures such as preference and perceived beauty would also help separate emotional valence from aesthetic liking.
Since prior research suggests that dimensional and discrete models capture different aspects of emotion and may rely on different processes, combining these approaches could provide a more accurate and nuanced understanding of emotional responses to music.

**Limitations**

Several limitations should be noted.
First, the sample size may have been too small to detect subtle effects.
We also did not recruit participants across the full range of musical sophistication, which likely reduced variability and made it harder to observe any moderating effects.
Using only Western music clips may limit emotional responses and reduce the generalizability of the findings across cultures.
In addition, the laboratory setting was highly controlled and may not reflect real-world musical experiences.
Our sample consisted mainly of university students in their early twenties, which limits generalizability to other age groups.
We did not analyze Brainsight data to confirm accurate targeting of M1.
We also did not verify that cortical inhibition was maintained after the music listening task.
These factors may have reduced our ability to detect meaningful effects.

**Conclusion**

In conclusion, our findings challenge prior research suggesting that M1 plays a direct role in emotion production, as we observed no causal relationship.
Emotional ratings were not influenced by musical sophistication under M1 inhibition.
However, participants with higher Active and Emotional Engagement with music showed greater sensitivity to arousal.
These results help clarify which neural systems are involved in emotional processing.
This is important for guiding future research, as it suggests that targeting M1 alone may not be effective for influencing emotion.
Instead, it highlights the need to focus on broader networks when developing interventions for motor rehabilitation and the treatment of affective disorders.

# References

::: {#refs}
:::

## Plots

```{r Library, echo=FALSE, message=FALSE, warning=FALSE}
data <- read.csv("C:/Users/knopf/OneDrive/Documents/2026-04-09_BIOL430G_mini-project/2026-04-09_BIOLG_430G_mini-project/data/processed/music_task_data_participant-level.csv")

library(here)
library(ez)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(effectsize)
library(renv)
```

```{r Figure 2, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Note: Within-subject line plot depicting individual participants’ mean valence ratings across stimulation conditions (active vs. sham), with separate panels for each stimulus type. Each line represents a single participant, illustrating the direction and magnitude of change between conditions."}
##Figure 2
#Look at participant individual change between active and sham for meanValenceRating
ggplot(data, aes(x = condition, y = meanValenceRating, group = participant)) +
  geom_point() +
  geom_line(alpha = 0.5) +
  facet_wrap(~stimuliType) +
  labs(title = "Individual Differences in Valence Ratings by Condition and Stimulus Type") +
  theme_minimal()
```

```{r Figure 3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Note: Within-subject line plot depicting individual participants’ mean valence ratings across stimulation conditions (active vs. sham), with separate panels for each stimulus type. Each line represents a single participant, illustrating the direction and magnitude of change between conditions."}

##Figure 3
#Look at participant individual change between active and sham for meanArousalRating
ggplot(data, aes(x = condition, y = meanArousalRating, group = participant)) +
  geom_point() +
  geom_line(alpha = 0.5) +
  facet_wrap(~stimuliType) +
  labs(title = "Individual Differences in Arousal Ratings by Condition and Stimulus Type") +
  theme_minimal()
```

```{r Figure 4, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Note.Within-subject line plot depicting individual participants’ mean arousal ratings across stimulation conditions (active vs. sham), with separate panels for each stimulus type. Each line represents a single participant, illustrating the direction and magnitude of change between conditions."}
##Figure 4
ggplot(data, aes(x = GoldMSI_ActiveEngagement_63,
                 y = meanArousalRating,
                 color = stimuliType)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "GoldMSI Active Engagement Predicting Arousal Ratings",
    x = "GoldMSI Active Engagement",
    y = "Mean Arousal Rating",
    color = "Stimulus Type"
  ) +
  theme_minimal()
```

```{r Figure 5, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Note: Scatterplot illustrating the relationship between Gold-MSI Active Engagement scores and mean arousal ratings across stimulus types. Lines represent fitted linear regression slopes for each stimulus category, with shaded areas indicating 95% confidence intervals."}
##Figure 5
#plot 
ggplot(data, aes(x = GoldMSI_Emotions_42,
                 y = meanArousalRating,
                 color = stimuliType)) +
  geom_point(alpha = 0.45, size = 2) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Emotions and Arousal Ratings",
    x = "GoldMSI Emotional Engagement",
    y = "Mean Arousal Rating",
    color = "Stimulus Type"
  ) +
  theme_minimal(base_size = 13)
```

